Towards a Cognitive System That Can Recognize Spatial Regions Based on Context

### 2012-01-24 code/scenarios/cdsr/trunk/papers/AAAI2012/main-aaai12-cdsr.tex
editing eval section

The ability to understand and reason about \textit{spatial regions} is essential for cognitive systems performing tasks for humans in everyday environments. Some regions, such as whole rooms and corridors, are defined by clearly perceivable boundaries (e.g. walls and doors). However, many regions to which humans routinely refer are not so easily defined. Consider, for example, the aforementioned region \textit{the front of the classroom}. This region is not easily perceivable using just the geometry of the environment. Instead, it is defined by the objects present in the room (chairs, a desk, a chalkboard), their role in this context (seats for students to watch a teacher who writes on the chalkboard) and their configuration in space (the seats point toward the chalkboard). In this paper, we will refer to such regions as \textit{context-dependent spatial regions} (CDSRs).

% nah: These are the main claims from the ACS paper, but they can be reduced/dropped for aaai

Representing and reasoning about context-dependent spatial regions is beyond the capabilities of current cognitive systems, yet it is an important ability for three reasons:

\begin{itemize}

\item{If cognitive systems are to collaborate with humans in everyday environments, then they must be able to represent and reason about the spatial regions humans refer to. Many regions are best defined in a context-dependent manner, for example, a kitchen in a studio apartment, an aisle in a church or store, behind enemy lines in a military engagement, etc.}

\item{Cognitive systems must integrate different types of information, including geometric, semantic, and functional knowledge. The identification of context-dependent spatial regions provides a suitable task for studying this integration. 
% Only through such integration could a system determine, for example, whether a banquet hall has a dancefloor depending on the arrangement of the room.
}

\item{Cognitive systems must adapt their knowledge and experience to new situations. While humans do not demonstrate mastery after seeing a single example of a new concept, they are able to begin using it immediately, whilst incrementally refining it over time (cf. one-shot learning~\cite{Fei-Fei/etal:2006}). This has important implications with respect to the amount of knowledge engineering and training time required to create a cognitive system. For example, having trained a mobile robot to understand the front of one classroom, it is desirable for it to be able to identify the front of other classrooms and similar areas (e.g. theaters) without additional tutoring.}
\end{itemize}

% nah: removed large motivating picture, hoping we can make do with just our result pictures, or replace it with images used for ground truth

This paper presents a novel integration of technologies which allows an artificial cognitive system (specifically a mobile robot) to represent and reason about CDSRs. Our approach is founded on two assumptions. The first assumption is that CDSRs can be defined using qualitative spatial representations (QSRs) extracted from the sensor data of the system~\cite{Cohn:2001}. The second assumption is that semantically similar areas (e.g. two different classrooms) will feature similar CDSRs, and that these similarities can be recognised through analogy. The rest of the paper is structured following these assumptions. Section~\ref{sec:qsr-gen} describes how we generate QSRs from sensor data taken from an existing, state-of-the-art, cognitive system and use these to define CDSRs. Section~\ref{sec:analogy} then describes how use the structure-mapping model of analogy \cite{Gentner1983a} to transfer a CDSR from a labelled example to a new situation. Section~\ref{sec:example} presents a worked example of the entire process, and Section~\ref{sec:evaluation} evaluates our system in comparison to data from human subjects performing the same task.

\section{From Sensors to Qualitative Representations}\label{sec:qsr-gen}

@@

\begin{figure}[h]
  \includegraphics[width=\columnwidth]{analogy2.png}
  \caption{Analogical mapping between four base expressions and two target expressions. Ovals and squares represent relations and entities respectively. Purple dashed bi-directional arrows are the correspondences between the base and target, and the green dashed arrow, oval, and square represent a candidate inference projected into the target.}
  \label{fig:analogy}
\end{figure}

>>>>

The ability to understand and reason about \textit{spatial regions} is essential for cognitive systems performing tasks for humans in everyday environments. Some regions, such as whole rooms and corridors, are defined by clearly perceivable boundaries (e.g. walls and doors). However, many regions to which humans routinely refer are not so easily defined. Consider, for example, the aforementioned region \textit{the front of the classroom}. This region is not easily perceivable using just the geometry of the environment. Instead, it is defined by the objects present in the room (chairs, a desk, a whiteboard), their role in this context (seats for students to watch a teacher who writes on the whiteboard) and their configuration in space (the seats point toward the whiteboard). We refer to such regions as \textit{context-dependent spatial regions} (CDSRs).

Current cognitive systems are not capable of representing and reasoning about CDSRs, yet it is an important ability. If cognitive systems are to collaborate with humans in everyday environments then they must be able to understand the spatial regions humans refer to. Many regions are best defined in a context-dependent manner, for example, a kitchen in a studio apartment, an aisle in a church or store, behind enemy lines in a military engagement, etc. In order to represent and reason about such regions, cognitive systems must integrate different types of information, including geometric, semantic, and functional knowledge. Creating systems able to integrate such a range of information is a key challenge in the cognitive systems paradigm.

This paper presents an artificial cognitive system (specifically a mobile robot) able to represent and reason about CDSRs. Our approach is founded on two assumptions. The first assumption is that CDSRs can be defined using \emph{qualitative spatial representations} (QSRs) extracted from the sensor data of the system~\cite{Cohn:2001}. The second assumption is that semantically similar areas (e.g. two different classrooms) will feature similar CDSRs, and that these similarities can be recognised through \emph{analogy}. The rest of the paper is structured following these assumptions. Section~\ref{sec:qsr-gen} describes how we generate QSRs from sensor data taken from an existing, state-of-the-art, cognitive system and use these to define CDSRs. Section~\ref{sec:analogy} then describes how we use the structure-mapping model of analogy \cite{Gentner1983a} to transfer a CDSR from a labelled example to a new situation. Section~\ref{sec:example} presents a worked example of the entire process, and Section~\ref{sec:evaluation} evaluates our system in comparison to data from human subjects performing the same task.

\section{Metric to Qualitative Representations}\label{sec:qsr-gen}

@@

\begin{figure}[h]
  \includegraphics[width=\columnwidth]{images/AnalogyDiagramv4.png}
  \caption{Analogical mapping between four base expressions and two target expressions. Ovals and squares represent relations and entities respectively. Purple dashed bi-directional arrows are the correspondences between the base and target, and the green dashed arrow, oval, and square represent a candidate inference projected into the target.}
  \label{fig:analogy}
\end{figure}
